|

### 2024-08-21 reproducibility/manuscript/manuscript.qmd
docs(manuscript): add posterior sampling methods
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

@@

@@

```python
scv.pp.filter_and_normalize(
  adata=adata,
  min_shared_counts=30, 
  n_top_genes=2000,
)
scv.pp.moments(
  adata,
  n_pcs=30,
  n_neighbors=30,
)
```

@@

```python
scv.pp.filter_and_normalize(
  adata=adata,
  n_top_genes=2000,
  min_shared_counts=20,
)
scv.pp.moments(adata)
```

>>>>

toc-depth: 2

@@

## Posterior prediction {#sec-methods-posterior-prediction}

To benchmark Pyro&thinsp;-Velocity performance in predicting cell fate, we
generated the posterior samples measurement
$x_{cg}=\left(u\left(\tau^{(k_{cg})}\right),
s\left(\tau^{(k_{cg})}\right)\right)$ or $x_{cg}=\left(u_{cg}, s_{cg} \right)$,
$t_c$, $\beta_g$, $\gamma_g$ from the same single cell RNA-seq using $N=30$
Monte Carlo samples from the posterior predictive distribution following
$p(x_{cg} \vert \theta, \psi) p(\theta, \psi) \approx p(x_{cg} \vert \theta,
\psi) q_{\phi}(\theta, \psi)$. $x_{cg}$ can be posterior samples of either
denoised gene expression (used for phase portraits and vector field-based
trajectory inference) or raw read counts (used for gene selection). Then we
calculated the posterior samples of RNA velocity as
$v_{cg}=\frac{ds(\tau^{(k_{cg})})}{d \tau^{(k{cg})}}$ based on posterior samples
measurement of denoised gene expression $x_{cg}$ and $\beta_{g}$, $\gamma_{g}$.

## Prioritization of cell fate markers {#sec-methods-fate-markers}

We prioritize the cell fate markers using two metrics: First, the Pearson
correlation between each gene's posterior mean of the denoised spliced
expression and posterior mean of the shared time; Second, the negative mean
absolute errors of each gene's observed spliced, and unspliced read counts with
posterior predictive samples of spliced and unspliced raw read counts, i.e.
$\frac{1}{N n_c} \sum_i \sum_c (x_{icg}-\tilde{X}_{cg})$, where $N$ is the
posterior sample number that is set to $30$, $i$ is the posterior sample index,
and $n_c$ is the cell number. We first select the top $300$ genes with the
highest negative mean absolute errors and then rank the $300$ genes based on the
most positively correlated genes and the least negatively correlated genes. We
use the same strategy for scVelo to rank the markers by the model likelihood to
get the top 300 genes and then prioritize these genes by Pearson's correlation
between scVelo latent time and normalized expression.

@@

```python
import scvelo

scvelo.pp.filter_and_normalize(
  adata=adata,
  min_shared_counts=30, 
  n_top_genes=2000,
)
scvelo.pp.moments(
  adata,
  n_pcs=30,
  n_neighbors=30,
)
```

@@

```python
scvelo.pp.filter_and_normalize(
  adata=adata,
  n_top_genes=2000,
  min_shared_counts=20,
)
scvelo.pp.moments(adata)
```
